---
title: "R Homework 1"
subtitle: "Instructions"
author: "Kyle Butts"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: ['default']
    ## No title slide
    seal: false
    self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r preamble, child=here::here("document-preamble.Rmd")}
```


```{r, echo = F}
library(icons)
library(tidyverse)
library(gt)

```

## Sample Means

Consider observing two years worth of sales at your company at a monthly interval (\\(n = 24\\)). Monthly sales have a mean of $1000 and a standard deviation of 100. 

- What is the distribution of the sample mean? Sketch this distribution.

<div style="margin-top: 160px"></div>
- What is the probability that a given sample of 24 months has a sample mean of $990 or less?


<div style="margin-top: 160px"></div>
- Now, assume you have two more years of data (\\(n = 48\\)). What is the probability that a given sample of 24 months has a sample mean of $990 or less? Why is this probability different?


---
## Confidence Intervals

Consider observing two years worth of sales at your company at a monthly interval (\\(n = 24\\)). Monthly sales have a mean of $1000 and a standard deviation of 100. 

- Using the true population mean of $1000, find the middle 90% of sample means.



<div style="margin-top: 160px"></div>
Now, you don't know the population mean and instead just have that the sample mean is $1020. 

- Create a confidence interval for this estimate and circle the margin of error. 

- Is it true that 90% of sample means falls within your confidence interval? Why not?

- Interpret the confidence interval correctly. 




---
## Hypothesis Testing

Consider observing two years worth of sales at your company at a monthly interval (\\(n = 24\\)). Monthly sales have a standard deviation of 100. Our sample has a mean of $1020. Consider the following test:

\begin{align*}
	H_0: \mu &= 1000 \\
	H_A: \mu &\neq 1000 \\
\end{align*}

- Sketch the sample distribution under the null and on the sketch, show what the p-value is


<div style="margin-top: 160px"></div>
- Test the hypothesis with a level of significance $\alpha = 0.10$. Do you reject the null?





<div style="margin-top: 160px"></div>
- Without redoing work (just think about it), how would the p-value you calculated above compare to the p-value from this test:

\begin{align*}
	H_0: \mu &= 1000 \\
	H_A: \mu &> 1000 \\
\end{align*}






---
## Rejection Region

Consider observing two years worth of sales at your company at a monthly interval (\\(n = 24\\)). Monthly sales have a standard deviation of 100. Consider the following test:

\begin{align*}
	H_0: \mu &= 1000 \\
	H_A: \mu &\neq 1000 \\
\end{align*}

- Calculate the rejection region with a level of significance $\alpha = 0.10$. Use this rejection region to determine whether you would reject the null with a sample mean of 1020.




<div style="margin-top: 160px"></div>
- Draw this rejection region. Compare this to the middle 90% of sample means with a population mean of 1000 that you calculated on the first page. They should be highly related!!


