---
title: "R Homework 5"
subtitle: "Instructions"
author: "Kyle Butts"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: ['default']
    ## No title slide
    seal: false
    self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r preamble, child=here::here("document-preamble.Rmd")}
```


```{r, echo = F}
library(icons)
library(tidyverse)
library(gt)

output_format <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

<h1 class="hi text-2xl center coral">Instructions for R Homework 5*</h1>

<div class="footnote"><sup>*</sup> For all HW assignments, I need to see all the code used</div>




## Question 1 
For this exercise we will run a regression using Swiss demographic data from around 1888. The sample is a cross-section of French speaking counties in Switzerland. This data come with the R package `datasets`. To load this dataset type this `data(swiss, package="datasets")`. 

The basic variable definitions are as follows:

```{r fertility-info, echo = F}
fertility <- tribble(
  ~Variable, ~Description,
  "Fertility",	"Ig, ‘common standardized fertility measure’",
  "Agriculture",	"% of males involved in agriculture as occupation",
  "Examination",	"% draftees receiving highest mark on army examination",
  "Education",	"% education beyond primary school for draftees.",
  "Catholic",	"% ‘catholic’ (as opposed to ‘protestant’).",
  "Infant.Mortality",	"live births who live less than 1 year.",
)
gt::gt(fertility) %>% 
  gt::tab_options(table.font.size = px(10))

```


Type `help(swiss)` in the console for additional details. Use the `summary()` command to report the mean and median for the variables Fertility, Education, and Catholic.


---
## Question 2 

We want to estimate the expected Fertility level in a Swiss county conditional on the county’s education level. We assume the relationship is linear. So, we are interested in estimating $\alpha$ and $\beta$ in 

$$
	\text{Fertility}_c = \alpha + \beta \cdot \text{Education}_c + \epsilon_c.
$$ 

If we use Ordinary Least Squares to estimate and we have the following formulas: 

$$ 
	\hat{\beta} = r_{x,y} \frac{s_y}{s_x}
$$

$$
	\hat{\alpha} = \bar{y} - \hat{\beta} \bar{x},
$$

where $y$ is the left hand side variable, $x$ is the right hand side variable, and the bar $\bar{ }$  denotes the sample mean, $s$ is the standard deviation, and $r_{x,y}$ is the correlation between $x$ and $y$.

- Find the correlation between Education and Fertility using the `cor()` function, as well as the sample standard deviation for each variable using the `sd()` function. Report these numbers.

- Use the `cor()` and `sd()` function to get an estimate for $\beta$ in the equation relating Fertility to Education. Keep this value stored in a scalar called `beta_hat`. Report this number by having a line with `print(beta_hat)`.

- Now use the estimate `beta_hat`, along with the function `mean()` to get an estimate for `alpha`. Keep this value stored as a scalar called `alpha_hat`. Report this number by having a line with `print(alpha_hat)`.



## Question 3 

Use `alpha_hat` and `beta_hat` to predict the average fertility rate in a county where 40% of the population is educated.


## Question 4 

Plot the relationship between Fertility and Education using the `plot()` function with Education on the horizontal axis. Make sure to label your axis! (If you don't know how, use `?plot`).


---

## Question 5 

Now estimate the model the model relating Fertility Rate to Education using the `lm()` function in R’s base code. Typically, if you want to estimate you use the syntax `lm(yvar ~ xvar, data = dataframe)`.

- Store the estimation results as follows `model_1 <- lm(...)`. This list should include a number of details include the estimated parameters, the coefficient of determination (r-squared), all of the residuals from the model, and more.

- Use the command `summary(model_1)` to report the summary of the ordinary least squares estimation. Do you have the same estimates as Question 2?

- What is the R-squared from this regression? Interpret it in a meaningful way.



## Question 6 

For each one of the estimated parameters reported in Question 5:

- Interpret the coefficient in a meaningful way.

- Report the results from testing the null hypothesis that the true parameter value is zero.

## Question 7 
Recreate the figure in Question 4, and then add the line of best fit using the `abline()` function with the coefficients from `model_1`, `model_1$coefficients`.

## Question 8

Use the `mean()` command to show that the average of the residuals associated with model_1 is zero. 
