---
title: "ECON 3818"
subtitle: "Distributions"
author: "Kyle Butts"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: ['default']
    self_contained: true
nature:
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
---
exclude: true

```{R, setup, include = F}

library(tidyverse)
library(svglite)
library(knitr)
library(here)
library(gt)
library(patchwork)
library(kfbmisc)

# Knitr options
opts_chunk$set(
comment = "#>",
fig.align = "center",
fig.height = 7,
fig.width = 10.5,
warning = F,
message = F
)

# Use svg for figures
opts_chunk$set(dev = "svglite")
options(device = function(file, width, height) {
svglite::svglite(tempfile(), width = width, height = height)
})

# HTML output for knitr
options(knitr.table.format = "html")


make_gt_title <- function(title) {
gt::html(glue::glue("<span class='hi slate' style='display: block; margin-bottom: 8px;'>{title}</span>"))
}

xaringanExtra::use_scribble()
```

---
class: clear, middle
<!-- Custom css -->
```{css, echo = F, code=xfun::read_utf8(here::here("Lecture Slides", "my-css.css"))}
```


<!-- From xaringancolor -->

<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{purple}{rgb}{0.337254901960784, 0.00392156862745098, 0.643137254901961}$$
$$\require{color}\definecolor{navy}{rgb}{0.0509803921568627, 0.23921568627451, 0.337254901960784}$$
$$\require{color}\definecolor{ruby}{rgb}{0.603921568627451, 0.145098039215686, 0.0823529411764706}$$
$$\require{color}\definecolor{alice}{rgb}{0.0627450980392157, 0.470588235294118, 0.584313725490196}$$
$$\require{color}\definecolor{daisy}{rgb}{0.92156862745098, 0.788235294117647, 0.266666666666667}$$
$$\require{color}\definecolor{coral}{rgb}{0.949019607843137, 0.427450980392157, 0.129411764705882}$$
$$\require{color}\definecolor{kelly}{rgb}{0.509803921568627, 0.576470588235294, 0.337254901960784}$$
$$\require{color}\definecolor{jet}{rgb}{0.0745098039215686, 0.0823529411764706, 0.0862745098039216}$$
$$\require{color}\definecolor{asher}{rgb}{0.333333333333333, 0.372549019607843, 0.380392156862745}$$
$$\require{color}\definecolor{slate}{rgb}{0.192156862745098, 0.309803921568627, 0.309803921568627}$$
$$\require{color}\definecolor{cranberry}{rgb}{0.901960784313726, 0.254901960784314, 0.450980392156863}$$
</div>
	
<script type="text/x-mathjax-config">
	MathJax.Hub.Config({
		TeX: {
			Macros: {
				purple: ["{\\color{purple}{#1}}", 1],
				navy: ["{\\color{navy}{#1}}", 1],
				ruby: ["{\\color{ruby}{#1}}", 1],
				alice: ["{\\color{alice}{#1}}", 1],
				daisy: ["{\\color{daisy}{#1}}", 1],
				coral: ["{\\color{coral}{#1}}", 1],
				kelly: ["{\\color{kelly}{#1}}", 1],
				jet: ["{\\color{jet}{#1}}", 1],
				asher: ["{\\color{asher}{#1}}", 1],
				slate: ["{\\color{slate}{#1}}", 1],
				cranberry: ["{\\color{cranberry}{#1}}", 1]
			},
			loader: {load: ['[tex]/color']},
			tex: {packages: {'[+]': ['color']}}
		}
	});
</script>

<style>
	.purple {color: #5601A4;}
	.navy {color: #0D3D56;}
	.ruby {color: #9A2515;}
	.alice {color: #107895;}
	.daisy {color: #EBC944;}
	.coral {color: #F26D21;}
	.kelly {color: #829356;}
	.jet {color: #131516;}
	.asher {color: #555F61;}
	.slate {color: #314F4F;}
	.cranberry {color: #E64173;}
</style>


```{r flair_color, echo=FALSE}
	library(flair)
	purple <- "#5601A4"
	navy <- "#0D3D56"
	ruby <- "#9A2515"
	alice <- "#107895"
	daisy <- "#EBC944"
	coral <- "#F26D21"
	kelly <- "#829356"
	jet <- "#131516"
	asher <- "#555F61"
	slate <- "#314F4F"
	cranberry <- "#E64173"
```


## Distributions

---
# Outline

Discrete Case

- Probability Mass Function
- Calculating probabilities


Continuous Case

- Probability Density Function
- Calculating probabilities





---
# Probability Distribution Functions

Until now we have discussed probability distributions in very loose terms. We will build a formal definition of a .hi.cranberry[probability distribution function].


First we consider the discrete case, and then the continuous case.


---
class: clear,middle

### Discrete

---
# Defining Probability Mass Function
Let $X$ be a discrete random variable defined over sample space $S$ with outcome $x \in S$.


The .hi.cranberry[probability mass function] (or pmf) of $X$ is a function that assigns a probability value to every possible outcome of $X$. 

We write this as $P_X(x)$, probability of $X=x$.


---
# Example PMF
.subheader.coral[Explicitly Given]

$X$ is defined as the number of people seated at a random table at a restaurant. The PMF of X is provided below:

```{r example-pmf, echo=F}

pmf <- tribble(
  ~x, ~`P(X = x)`,
  "1", 0.07,
  "2", 0.36,
  "3", 0.32,
  "4", 0.21,
  "5 or more", 0.04,
)

pmf_tbl <- gt(pmf) %>% 
  gt_theme_kyle() %>%
  tab_header(title = "Probability Distribution of X")

pmf_tbl

```



---
# Example PMF
.subheader.coral[Based on Scenario]

Suppose you flip a fair coin twice. Let $X$ be the number of heads that appear. The pmf of $X$ is

```{r pmf-coin, echo=F}
coin_pmf <- tribble(
  ~x, ~`P_X(x)`,
  "0", 1/4,
  "1", 1/2,
  "2", 1/4,
)

gt(coin_pmf) %>% 
  gt_theme_kyle() %>%
  tab_header(title = "Probability Distribution of X")
```



---
# Example PMF

```{r, echo = F, out.width = "90%"}
ggplot(coin_pmf) + 
  geom_bar(aes(x = x, weight = `P_X(x)`), stat="count") +
  labs(y = "P_X(x)", x = "X") +
  theme_kyle(base_size = 20)
```


---
# Properties of PMFs

We say a $p_X(x)$ is a .hi.purple[valid] pmf if it satisfies the following:

1. $0 \leq p_X(x) \leq 1$ for all $x \in S$.

2. $\displaystyle\sum_{x \in S} p_X(x) = 1.$



---
# Using PMFs

We can use the PMF to answer questions about cumulative probabilities, for example:
Recall the previous example:


`r pmf_tbl`


What is the probability a random table at the restaurant has 2 or 3 people seated?
$$ 
P(X=2) = 0.36 \text{ and } P(X=3) = 0.32 \implies 
$$

$$
P(X=2 \text{ or } 3) = 0.36 + 0.32 = .68 
$$



---
# Clicker Question
Assume there are four outcomes of $X$: $1, 5, 10$ and $20$. Given the following PMF, what is the probability X=20?

```{r pmf-missing, echo=F}

pmf_missing <- tribble(
  ~x, ~`P(X = x)`,
  "1", "0.42",
  "5", "0.23",
  "10", "0.18",
  "20", "?",
)

gt(pmf_missing) %>% 
  gt_theme_kyle() %>%
  tab_header(title = "Probability Distribution of X")

```

<ol type = "a">
  <li>0.35</li>
  <li>0.17</li>
  <li>0.40</li>
  <li>Cannot be determined given the information</li>
</ol>


---
class: clear,middle

### Continuous


---
# Defining Probability Density Function
Let $Y$ be a continuous random variable defined over the interval $[a,b]$.


The .hi.purple[probability density function] (or pdf) of $Y$ is a function, $f_Y(y)$, that assigns a probability value to every possible .it[interval] in $[a,b]$. 

We write
$$ Pr(c\leqslant Y\leqslant d) = \int_{c}^{d}f_Y(y) dy, $$
for all $(c,d)\subset [a,b]$.

---
# Example pdf

For $Z \sim N(0,1)$, find $P(Z <= -1)$.


```{r, echo = F, out.width = "100%"}

bar <- tibble(xmin = seq(-4, -1.2, by = 0.5), xmax = xmin + 0.5, ymin = 0, ymax = dnorm((xmin + xmax)/2)) 

p2 <- ggplot() +
    geom_rect(
        data = bar, 
        aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
        fill = "#d2382c", alpha = 0.5) + 
    # stat_function(
    #     data = data.frame(x = c(-4, -1)),
    #     fun = dnorm,
    #     xlim = c(-4, -1), size = 0,
    #     geom = "area", fill = "#d2382c", alpha = 0.5
    # ) +
    # Normal curve
    stat_function(
        data = data.frame(x = c(-4, 4)), 
        fun = dnorm, xlim = c(-4, 4),
        color = "black", size = 1.5) +
    scale_x_continuous(breaks = -4:4) +
    theme_kyle(base_size = 20) +
    labs(x = "Value", y = "Density", title = "PMF-approximation")

p3 <- ggplot(data.frame(x = c(-4, 4)), aes(.data$x)) +
    stat_function(
        fun = dnorm,
        xlim = c(-4, -1), size = 0,
        geom = "area", fill = "#d2382c", alpha = 0.5
    ) +
    # Normal curve
    stat_function(fun = dnorm, color = "black", size = 1.5) +
    scale_x_continuous(breaks = -4:4) +
    theme_kyle(base_size = 20) +
    labs(x = "Value", y = "Density", title = "True area")

```

```{r, echo = F, out.width = "80%"}
print(p2)
```


---
# Integral of PDF = Probability


```{r, echo = F, out.width = "100%"}
p2 + p3
```


---
# Example PDF
Suppose that $Y$ is a continuous random variable with pdf $f_Y(y) = 3y^2$ for $0<y<1$. What is $P(\frac{1}{4} \leqslant Y \leqslant \frac{1}{2})$?

```{r, echo = F, out.width = "90%"}

ex_pdf <- function(y) {3*y^2*(y > 0 & y < 1)}


ggplot(data.frame(x = c(-1, 2)), aes(.data$x)) +
    stat_function(
        fun = ex_pdf,
        xlim = c(1/4, 1/2), size = 0,
        geom = "area", fill = "#d2382c", alpha = 0.5
    ) +
    stat_function(fun = ex_pdf, color = "black", size = 1.5) +
    scale_x_continuous(breaks = -1:2) +
    theme_kyle(base_size = 20) +
    labs(x = "Value", y = "Density")

```


---
# Example PDF
Suppose that $Y$ is a continuous random variable with pdf $f_Y(y) = 3y^2$ for $0<y<1$. What is $P(\frac{1}{4} \leqslant Y \leqslant \frac{1}{2})$?


---
# Properties of PDFs
We say a $f_Y(y)$ is a .hi.purple[valid] pdf if it satisfies the following:


1. $0\leqslant \int f_Y(y) \leqslant 1$ for all $y \in [a,b]$.

2. $\displaystyle\int_{a}^b f_Y(y)dy = 1.$


Note that $Pr(Y=a) = \displaystyle\int_{a}^a f_Y(y)dy = 0$. 

At first this might seem counterintuitive. But imagine trying to stop a stopwatch at exactly 30 seconds. What is the probability of that event?


---
# Clicker Question
Given the pdf, $f(y)=3y^2$ for $0<y<1$. What is the $P(Y<1/3)$?

<ol type = "a">
  <li>\(\frac{1}{3}\)</li>
  <li>\(\frac{1}{9}\)</li>
  <li>\(\frac{1}{27}\)</li>
  <li>\(\frac{26}{27}\)</li>
</ol>


---
# Midterm Example
Consider the probability distribution for random variable Y:

$$ 
f(y)= 8y, \ 0 \leq y \leq \frac{1}{2} 
$$

1. Find $P(Y<\frac{1}{3})$
2. Find $P(Y=\frac{1}{4})$
3. Find $P(\frac{3}{4}<Y<1)$

