---
title: "R Day 5 - Size and Power"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# Size and Power

## Size

Let $X \sim N(10, 4^2)$

Let's see how often we reject the null when the **null is the true population mean** (i.e. size of our test). We will use the level of confidence $\alpha = 0.05$.

$$ H_0: \mu = 10$$ 
$$ H_A: \mu \neq 10 $$
Let's calculate the power to detect a true population mean of $\mu = 11$. To do so, we are going to do the following:

1. Simulate data using the null mean of 10

2. Calculate a z-score and a two-tailed $p$-value

3. Track in a vector if we reject the null hyptohesis

4. Repeat steps 1-3 `n_sims` times

Then, we will count the number of times we (incorrectly!) reject null


```{r "Size"}
# Set Parameters
n <- 50
true_mean <- 10
null_mean <- 10
true_sd <- 4
alpha <- 0.10


```


**Exercise:**

- Change the standard deviation. How does the size change?

- Change the number of observations $n$. How does the size change? 
  - Why not? I thought more observations gave us more precise estimates?


## Power

Let's see how often we reject the null when the true population mean is different from our null (i.e. power of our test). We will use the level of confidence $\alpha = 0.05$.
$$ H_0: \mu = 10$$

$$ H_A: \mu \neq 10 $$

- To calculate the power, first we need to calculate the rejection region

$$ 
	P(\bar{X} \geq X^*) = .025 \implies P(Z \geq \frac{X^* - 10}{4/\sqrt{50}}) = .025
$$ 

This yields a critical $Z$-score of $1.96$

$$ 
	\frac{X^* - 10}{4/\sqrt{50}} = 1.96 \implies X^* = 11.11, \text{ and }
$$ 

$$ 
	\frac{X^* - 10}{4/\sqrt{50}} = - 1.96 \implies X^* = 8.89
$$ 

Therefore, we reject the null whenever $X^* \geq 11.11$ or $X^* \leq 8.89$

- Then, power is the probability of being in the rejection region, given a true population mean, in our case $\mu = 11$. 

$$
	P(\bar{X} \geq 11.11 \vert \mu = 11) + P(\bar{X} \leq 8.89 \vert \mu = 11)
$$

$$
	P(Z \geq \frac{11.11 - 11}{4/\sqrt{50}}) = P(Z \geq 0.1944) = 0.422
$$

$$
	P(Z \leq \frac{8.89 - 11}{4/\sqrt{50}}) = P(Z \leq -3.73) \approx 0
$$

Therefore our calculated power is: 
$$
	P(\bar{X} \geq 11.11 \vert \mu = 11) + P(\bar{X} \leq 8.89 \vert \mu = 11) = 0.422
$$


```{r "Power"}
# Set Parameters
n <- 50
true_mean <- 11
null_mean <- 10
true_sd <- 4
alpha <- 0.10


```


**Exercise:**

- Change the standard deviation. How does the power change?

- Change the number of observations $n$. How does the power change?
 
- Change the true population mean. How does the power change?


