---
title: "R Day 6 - t-stat"
output:
  html_document:
    theme: spacelab
    highlight: tango
---


# Two-sided t-test

**Exercise:** fill in the missing code to calculate a two-sided t-test by hand.

$$ 
	H_0: \mu_1 - \mu_2 = 0 
$$

$$
	H_A: \mu_1 - \mu_2 > 0
$$

Our z-stat in this case is 

$$
	Z\text{-stat} = \frac{(\bar{X}_1 - \bar{X}_2) - (\mu_1 - \mu_2)}{\sqrt{\sigma_1^2/n_1 + \sigma_2^2/n_2}}
$$

And our p-value is 
$$
	P (Z \geq Z\text{-stat}) 
$$

```{r "two-sided Z-test"}
set.seed(390)

# Random Sample of group 1 and 2
x1 <- rnorm(50, mean = 10, sd = 5)
x2 <- rnorm(35, mean = 11, sd = 5)

# Calculate z-stat


# Calculate p-value


```


```{r "two-sided t-test"}
set.seed(14)

# Random Sample of group 1 and 2
x1 <- rnorm(24, mean = 10, sd = 5)
x2 <- rnorm(15, mean = 11, sd = 5)

# Calculate t-stat


# Calculate p-value
# Note you have to supply t_score and degrees of freedom to `pt()`

```


Let's compare that with R's built-in Student's t-test. Type `?t.test` to learn the syntax for calculating a $t$-test

```{r}
# ?t.test
```



**Exercise:**

$$H_0 : \mu_1 - \mu_2   =  0$$
$$H_1 : \mu_1 - \mu_2 \neq 0$$

```{r}
set.seed(14)

# Random Sample of group 1 and 2
x1 <- rnorm(24, mean = 10, sd = 5)
x2 <- rnorm(15, mean = 11, sd = 5)

t.test(x1, x2, alternative = "two.sided")
```

1. Holding fixed the means, what happens to the p-value when you increase the sample size of the two groups? 

2. Holding fixed the sample size, what happens to the p-value as you move the two population means away from eachother?




# t-test for car color

Let's try to answer the following question:

> Do red cars and black/white cars sell for the same price?

$$ H_0: \mu_r - \mu_{bw} = 0$$ 
$$ H_A : \mu_r - \mu_{bw} > 0$$

To do so, let's use the USA Cars dataset

```{r "Load Car Data"}
# Load USA_cars.csv


# Subset data


# 2 sample Z-test


```

Since the p-value is larger than 0.05, with 95% confidence, I fail to reject that red cars and black cars sell for the same price.
