---
title: "R Day 7 - Regression"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# Load Wage Data

```{r, include = FALSE}
library(wooldridge)
data(wage1)

head(wage1)
```

# Regression

```{r}
# Scatter plot of wage on education
```

```{r}
# Remove education outliers (educ <= 5)

```


# Regressions

Regression is done by `lm()` function. Specifically the format looks like this:

`reg <- lm(y_var ~ x_var, data = df)`

To see the summary, you can type `summary(reg)`.

```{r}
# wage = a + b * educ

```

```{r}
# Scatter Plot and add regression line using abline

```


**Exercise:**

1. How do you interpret the coefficient on `educ`?

2. Is the coefficient on education statistically significant? How do you know?




# Regression with Indicator Variables

Let's see if individuals earn more money from having a high-school degree. To do so, we have to create a HS dummy variable

```{r}
# Create a high school dummy
```

**Exercise:**

1. How do you interpret the coefficient on `hs`?

2. Let's call this the "degree premium". Find a confidence interval using `confint` for the degree premium. Interpret this

```{r}

```



# Simpson's Paradox

```{r Load data}
library(palmerpenguins)
data(penguins)

head(penguins)
```

```{r}
# Scatter plot of Bill Depth (mm) and Body Mass (g)

# Regression
```



```{r}
# Seperately by penguin species
library(ggplot2)

ggplot(penguins, aes(x = bill_depth_mm, y = body_mass_g)) + 
	geom_point(aes(color = species)) + 
	geom_smooth(method = "lm", aes(color = species))

```

















