---
title: "R Day 3 - Distributions"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# Distributions

R has a suite of distribution functions that will make life *WAY* easier. For each distribution you will have a corresponding set of for functions: `rxxxx`, `pxxxx`, `qxxxx`, and `dxxxx`, where `xxxx` can be, for example, `binom` or `norm` for the binomial and normal distributions respectively. Type `?Distributions` if you get confused.

- `rxxxx` draws random observations from the distribution function

   - You give it $n$, and it returns $n$ generated observations from the distribution

- `pxxxx` calculates the cumulative distribution function

   - You give it $x$, and it returns $P(X \leq x)$.

- `qxxxx` calculates the "quantile" function

   - You give it $p$, a probability between $0$ and $1$ and it returns the $x$ such that $P(X \leq x) = p$

- `dxxxx` calculates the density at a given value of $x$ 

   - You give it $x$, and it returns $P_X(x) = P(X = x)$ in the discrete case and $f_X(x)$ in the continuous case .




# Binomial Distribution

Let's work with the following example. Shaq takes 5 free throws with a probability of making it of 52%. $X$ is the number of successfully made freethrows. Fill in the following information:

$$X \sim B(n = ?, p = ?)$$


## Calculating Probability of Binomial Distribution


### "By hand"

First, let's try to calculate the following probabilities: 

- $P(X = 1)$ 

- $P(X = 3)$

- $P(X >= 4)$ 

by hand. The function `choose(n,k)` will help.

```{r Probability by Hand}
p <- 0.52

# Calculate P(X = 1)


# Calculate P(X = 3)


# Calculate P(X >= 4)


```

### Using `dbinom`

`dbinom` gives $P(X = x)$ for a specified value of $x$. Type `?dbinom` to figure out what parameters to use. Using `dbinom` calculate:

- $P(X = 1)$ 

- $P(X = 3)$

- $P(X >= 4)$ 

```{r}

# P(X = 1)

# P(X = 3)

# P(X >= 4)

```

### Using `pbinom`

`pbinom` gives $P(X <= x)$ for a specified value of $x$. Using `pbinom` calculate:

- $P(X = 1)$ 

- $P(X = 3)$

- $P(X >= 4)$ 

```{r}

# P(X = 1)


# P(X = 3)


# P(X >= 4)


# P(X >= 4)

```



## Generating Random Variables from Binomial Distribution

We want to draw 100 random observations from this distribution. That is, we want to make Shaq shoot 100 sets of 5 freethrows and record the number of successes. Type `?rbinom` and figure out how to generate 100 observations.

```{r Drawing from Binomial Distribution}

```


Let's plot a histogram of the sample distribution of 10 observations with the argument `probability = TRUE` and to make the breaks for each integer, use `breaks = -0.5:0.5`. Do the sample probabilities match $P(X = 1)$ and $P(X = 3)$? Why or Why Not?

```{r}

```

Repeat the exercise below and increase the number of observations. What happens to the sample probabilities as n gets larger?

```{r}

```

#### Practice

Try the histogram for different values of size and prob. Describe what happens to the mean, the variance/standard deviation, and the skewness as you change `size`? as you change `prob`?

$$ \mu = np $$

$$ \sigma = \sqrt(np(1-p))$$

```{r Different values of size}
large_sample <- rbinom(n = 10000, size = 1000, prob = 0.52)

hist(large_sample, breaks= -0.5:1000.5, probability = TRUE)
```


```{r Different values of Prob}
large_sample <- rbinom(n = 10000, size = 5, prob = 0.3)

hist(large_sample, breaks= -0.5:5.5, probability = TRUE)
```






# Normal Distribution

The weight of the [second best apple in the world](https://theappleist.tumblr.com/toplist), the Honey Crisp, is 91 grams with a standard deviation of 3 grams.

## Generating Random Variables from Normal Distribution

Let's visualize the data by drawing 10,000 observations from this distribution using `rnorm` and plotting them using `hist`. Verify that 99% of the data falls within 3 standard deviations of the mean.

```{r}

```

## Calculating Probabilities

Now, let's say I want to know the probability that a honey crisp apple weighs less than 86 grams. We will use the `pnorm` function to determine this. We will do it two ways:

1. Specify the mean and the standard deviation

```{r}

```

2. First, standardizing the probability and using the $Z$-score.

$$P(X < 86) = P(Z < \frac{86 - 91}{3})$$

```{r}

```

Now let's calculate the following probabilities:

- $P(X <= 91)$

- $P(X > 93)$

- $P(90 < X < 92)$ 

```{r}
# P(X <= 91)


# P(X > 93)


# P(90 < X < 92)$

```


## Determining Percentiles of the Normal Distribution

Using `qnorm`, determine what the 20th, 60th, and 95th percentile of apple weight is. Interpret these numbers

```{r}
# 20th Percentile


# 60th Percentile


# 95th Percentile

```

Now, using what you got for the 95th percentile, show that the probability of observing an apple that ways less than the 95th percentile is indeed 95%

```{r}

```



