---
title: "R Day 7 - Regression"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# Load Wage Data

```{r, include = FALSE}
library(wooldridge)
data(wage1)

head(wage1)
```

# Regression

```{r}
plot(wage1$educ, wage1$wage)
```

```{r}
# Remove education outliers
wage1 <- wage1[wage1$educ >= 6, ]

plot(wage1$educ, wage1$wage)
```


# Regressions

Regression is done by `lm()` function. Specifically the format looks like this:

`reg <- lm(y_var ~ x_var, data = df)`

To see the summary, you can type `summary(reg)`.

```{r}
# wage = a + b * educ
reg <- lm(wage ~ educ, data = wage1)


reg
summary(reg)

```

```{r}
plot(wage1$educ, wage1$wage, col= "black")
abline(reg, col = "red")
```


**Exercise:**

1. How do you interpret the coefficient on `educ`?

2. Is the coefficient on education statistically significant? How do you know?




# Regression with Indicator Variables

```{r}
# Create a high school dummy
wage1$hs <- as.numeric(wage1$educ >= 12)

reg <- lm(wage ~ hs, wage1)
summary(reg)
```

**Exercise:**

1. How do you interpret the coefficient on `hs`?

2. Let's call this the "degree premium". Find a confidence interval using `confint` for the degree premium. Interpret this

```{r}
confint(reg)
```



# Simpson's Paradox

```{r Load data}
library(palmerpenguins)
data(penguins)

head(penguins)
```

```{r}
# Scatter plot of Bill Depth (mm) and Body Mass (g)
plot(penguins$bill_depth_mm, penguins$body_mass_g)

# Regression
reg1 <- lm(penguins$body_mass_g ~ penguins$bill_depth_mm)

summary(reg1)

plot(penguins$bill_depth_mm, penguins$body_mass_g)
abline(reg1) 
```



```{r}
# Seperately by penguin species
library(ggplot2)

ggplot(penguins, aes(x = bill_depth_mm, y = body_mass_g)) + 
	geom_point(aes(color = species)) + 
	geom_smooth(method = "lm", aes(color = species))

```

















