---
title: "R Day 4 - Sampling Distribution"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# For loops

Loops in R allow you to repeat a task multiple times. For example, let's say you want to print out the numbers 1 through 10 on separate lines in the console. You could do this by brute force, but the point of programming is to make things EASIER!

```{r}

```

A for loop is set up as 

$$
	\texttt{for}(\underbrace{\texttt{i}}_{\text{iterator}} \texttt{in } \underbrace{1:10}_{\text{values }\texttt{i}\text{ cycles through}}) \texttt{\{ }\text{Code goes here... } \texttt{\}}
$$

```{r}

```




# Binomial - prob = 0.3

Let's generate 100 observations from the following distribution:

$$ X \sim B(10, 0.3)$$

```{r "Binomial p=0.3 - One Trial"}
# 100 observations of Uniform Random Variable


# Histogram of sample


# True mean = 10 * 0.3 = 3
# Calculate Sample mean

```

This sample of size $100$ had a sample mean number of succeses of `r mean(sample)`. Let's try this exercise many times:

1. For sample $1$, generate 100 observations from $X \sim B(10, 0.3)$ and calculate the sample mean $\bar{X}_1$ for that sample

2. Store this sample in a vector of sample_means

3. Repeat steps 1 and 2, 1000 times to generate a sample distribution of the sample mean.

Do this 1000 times. Ahha! Copying and Pasting can't even help here, but for loops can!

```{r "Binomial p=0.3 - Sampling Distribution"}

```


**Exercise:**

```{r "Binomial p=0.3 - Play around with n_obs and n_trials"}

n_obs <- 100
n_trials <- 1000

sample_means <- c()

for(i in 1:n_trials){
	sample <- rbinom(n = n_obs, size = 10, prob = 0.3)
	mean <- mean(sample)
	
	sample_means <- c(sample_means, mean)
}

hist(sample_means)

```

Play around with `n_obs` and `n_trials` and answer the following questions:

- What happens to the distribution as `n_trials` increase for a fixed `n_obs`? From what we know from class, what is the sampling distribution for a fixed `n_obs`?

- What happens to this distribution as `n_obs` increases for a fixed `n_trials`?



# Confidence Intervals

Let's find the middle 95% of our sample means:

```{r "Binomial p=0.3 - Confidence Intervals"}

# Calculate population mean and population sd


# Critical value for normal distribution


# Calculate margin of error


# Calculate upper and lower confidence interval


# Plot histogram for sample distribution and add vertical lines for confidence interval
# abline will make vertical lines for us


```

## Confidence Interval Intuition

Using a for loop, let's show that a 95% confidence interval around the sample mean will contain the population means 95% of the time. 

That is $$P(\mu \in [X_{lower}, X_{upper}]) = 0.95$$


```{r "Confidence Interval Example"}



```












# Binomial - prob = 0.9

Let's now repeat this exercise for 

$$X \sim B(10, 0.9)$$

```{r "Binomial p=0.9 - One Trial"}

```

```{r "Binomial p=0.9 - Sampling Distribution"}

```

```{r "Binomial p=0.9 - Play around with n_obs and n_trials"}

```

- What happens to the distributino as `n_trials` increase for a fixed `n_obs`? From what we know from class, what is the sampling distribution for a fixed `n_obs`?

- What happens to this distribution as `n_obs` increases for a fixed `n_trials`?







# Uniform

One thing I tried to emphasize in the course is that for almost any underlying variable $X$, the **sampling distribution** of $\bar{X}$ will be normally distributed. It does not matter the underlying distribution! Let's try this with the uniform distribution.

$$ 
X \sim U(0, 1)
$$

```{r "Uniform - One Trial"}
# 100 observations of Uniform Random Variable
sample <- runif(n = 100, min = 0, max = 1)

# True mean = 0.5
mean(sample)
```

```{r "Uniform - Sampling Distribution"}
sample_means <- c()

for(i in 1:1000){
	sample <- runif(n = 100, min = 0, max = 1)
	mean <- mean(sample)
	
	sample_means <- c(sample_means, mean)
}

hist(sample_means)
```

```{r "Uniform - Play around with n_obs and n_trials"}
n_obs <- 100
n_trials <- 1000

sample_means <- c()

for(i in 1:n_trials){
	sample <- runif(n = n_obs, min = 0, max = 1)
	mean <- mean(sample)
	
	sample_means <- c(sample_means, mean)
}

hist(sample_means)

```


# More Complex Distribution

This will even work with weirder distributions! For example, let's try a distribution similar to income in the US.


```{r "More Complex - One Trial"}
sample <- 40 + 10 * rweibull(n = 1000, shape = 0.85)

# Histogram of sample
hist(sample, xlab = "Income ($1000)")
```

```{r "More Complex - Sampling Distribution"}
sample_means <- c()

for(i in 1:10000){
	sample <- 40 + 10 * rweibull(n = 1000, shape = 0.85)
	mean <- mean(sample)
	
	sample_means <- c(sample_means, mean)
}

hist(sample_means)
```
