---
title: "R Day 6 - t-stat"
output:
  html_document:
    theme: spacelab
    highlight: tango
---


# Two-sided t-test

**Exercise:** fill in the missing code to calculate a two-sided t-test by hand.

$$ 
	H_0: \mu_1 - \mu_2 = 0 
$$

$$
	H_A: \mu_1 - \mu_2 > 0
$$

Our z-stat in this case is 

$$
	Z\text{-stat} = \frac{(\bar{X}_1 - \bar{X}_2) - (\mu_1 - \mu_2)}{\sqrt{\sigma_1^2/n_1 + \sigma_2^2/n_2}}
$$

And our p-value is 
$$
	P (Z \geq Z\text{-stat}) 
$$

```{r "two-sided Z-test"}
set.seed(390)

# Random Sample of group 1 and 2
x1 <- rnorm(50, mean = 10, sd = 5)
x2 <- rnorm(35, mean = 11, sd = 5)

# Calculate means
mean1 <- mean(x1)
mean2 <- mean(x2)

var1 <- var(x1)
var2 <- var(x2)

n1 <- length(x1)
n2 <- length(x2)

# Calculate z-stat
z_stat <- ((mean1 - mean2) - 0) / (sqrt(var1/n1 + var2/n2))

# Calculate p-value
1 - pnorm(z_stat)


```


```{r "two-sided t-test"}
set.seed(14)

# Random Sample of group 1 and 2
x1 <- rnorm(24, mean = 10, sd = 5)
x2 <- rnorm(15, mean = 11, sd = 5)

# Calculate means
mean1 <- mean(x1)
mean2 <- mean(x2)

var1 <- var(x1)
var2 <- var(x2)

n1 <- length(x1)
n2 <- length(x2)

# Calculate t-stat
t_stat <- ((mean1 - mean2) - 0) / (sqrt(var1/n1 + var2/n2))

# Calculate p-value
# Note you have to supply t_score and degrees of freedom to tnorm()
1 - pt(t_stat, df = 13)

```


Let's compare that with R's built-in Student's t-test. Type `?t.test` to learn the syntax for calculating a $t$-test

```{r}
t.test(x1, x2, alternative = "greater")
```

**Exercise:**

$$H_0 : \mu_1 - \mu_2   =  0$$
$$H_1 : \mu_1 - \mu_2 \neq 0$$

```{r}
set.seed(14)

# Random Sample of group 1 and 2
x1 <- rnorm(24, mean = 10, sd = 5)
x2 <- rnorm(15, mean = 11, sd = 5)

t.test(x1, x2, alternative = "two.sided")
```

1. Holding fixed the means, what happens to the p-value when you increase the sample size of the two groups? 

2. Holding fixed the sample size, what happens to the p-value as you move the two population means away from eachother?




# t-test for car color

Let's try to answer the following question:

> Do red cars and black/white cars sell for the same price?

$$ H_0: \mu_r - \mu_{bw} = 0$$ 
$$ H_A : \mu_r - \mu_{bw} > 0$$

To do so, let's use the USA Cars dataset

```{r "Load Car Data", include = FALSE}
df <- read.csv("USA_cars.csv")

df
```


```{r, include = FALSE}

is_red <- df$color == "red"
is_bw <- df$color %in% c("black", "white")

n_red <- sum(is_red)
n_bw <- sum(is_bw)

# df[rows you want, columns you want]
cars_bw <- df[is_bw, ]
cars_red <- df[is_red, ]

```


```{r}
# 2 sample Z-test
mu_bw <- mean(cars_bw$price)
mu_red <- mean(cars_red$price)

sd_bw <- sd(cars_bw$price)
sd_red <- sd(cars_red$price)

z_score <- (mu_red - mu_bw)/sqrt(sd_red^2/n_red + sd_bw^2/n_bw)

1 - pnorm(z_score)

```

Since the p-value is larger than 0.05, with 95% confidence, I fail to reject that red cars and black cars sell for the same price.
