---
title: "Distribution Review"
author: "Kyle F. Butts"
date: "9/16/2020"
output: pdf_document
---

# Generating Random Variables

In R, `rxxxx` will denote generating random variables (r for random variable). The exact parameters depends on the distribution. We will look at 
$$X \sim B(n, p) \text{ and } Y \sim N(\mu, \sigma^2)$$ 

```{r Random Binomial}
# n = Number of observations, size = Number of trials per observation, p = Probability of Success
rbinom(n = 10, size = 10, p = 0.5)
```

```{r Random Normal}
rnorm(n = 10, mean = 0, sd = 2^2)
```

## Visualizing Distributions

```{r Binomial Histogram}
x <- rbinom(n = 10, size = 10, p = 0.5)
hist(x)
```


```{r Normal Histogram}
y <- rnorm(n = 10, mean = 0, sd = 2^2)
hist(y)
```



## Randomness in Draws

As the number of observations increase, the distribution will better approximate the underlying population distribution. For example, we will create the same histograms with n = 1000 observations.

```{r Binomial Histogram 2}
x <- rbinom(n = 1000, size = 10, p = 0.5)
hist(x)
```


```{r Normal Histogram 2}
y <- rnorm(n = 1000, mean = 0, sd = 2^2)
hist(y)
```



# Probability Mass/Density Function

In R, `dxxxx` will return the probability mass/density function (d for density). If `xxxx` is a discrete random variable like a binomial random variable, it will return 
$$p_X(x) = P(X = x).$$

For a continuous random variable like the normal distribution, it will return the density 

$$f_Y(y).$$

```{r}
# P(X = 5) with size = 10 trials and a probability of success of p = 0.3.
dbinom(5, size = 10, p = 0.3)
```

```{r}
# Density at 0. This is NOT the probability that P(Y = 0).
dnorm(0, mean = 0, sd = 2^2)
```



# Cummulative Probabilities

In R, `pxxxx` will return the left-tail probability of observing a value $\leq$ a specified value (p for cummulative probability).

```{r, echo = FALSE}
library(tidyverse)
library(svglite)

# Custom Fonts for ggplot2
library(showtext)
showtext_auto()
## Loading Google fonts (https://fonts.google.com/)
font_add_google("Fira Sans", "fira_sans")

df <- tibble(
	x = seq(-4, 4, by= 0.01), 
	prob = dnorm(x, mean = 0, sd = 1^2),
	less_than = x < -1
)

ggplot() + 
	geom_line(data = df, mapping = aes(x = x, y = prob)) + 
	geom_area(data = df %>% filter(less_than == 1), 
		mapping = aes(x = x, y = prob), 
		fill = "steelalice"
	) +
	geom_vline(xintercept = -1) + 
	geom_text(
		data = data.frame(x = -2, y = 0.2, label = "P(Z <= -1)"),
		mapping = aes(x = x, y = y, label = label),
		size = 6, alpha = 1, inherit.aes = FALSE
	) + 
	labs(
		title = "Cummulative Probability that Z <= -1 ",
		subtitle = "Z ~ N(0,1)",
		x = "Z",
		y = "Density"
	) + 
	theme_grey(
		base_size = 14,
		base_family = "fira_sans"	
	) + 
	theme(
		legend.position="bottom", 
		legend.title.align = 0.5,
		# No Background Color
		legend.background = element_rect(fill= NA), 
		panel.background = element_rect(fill = NA, color = NA),
		plot.background=element_rect(fill = "#ECECEC", color = NA),
		# Center title
		plot.title= element_text(hjust= 0.5),
		plot.subtitle = element_text(hjust = 0.5, face = "italic")
	) 


```


```{r}
# Left Tail Probability, 
# P(X <= 5) with size = 10 trials and a probability of success of p = 0.3
pbinom(5, size = 10, p = 0.3)

# Right Tail Probability
# P(X >= 5) with size = 10 trials and a probability of success of p = 0.3
pbinom(5, size = 10, p = 0.3, lower.tail = FALSE)
```

```{r}
# P(Z <= 1)
pnorm(1, mean = 0, sd = 1^2)

# P(X <= 5) with mean = 10, sd = 2^2
pnorm(5, mean = 10, sd = 2^2)
```




# Finding Critical Value

In R, `qxxxx` is the inverse of the `pxxxx`. It asks the question, for what value, $z^*$, will $p\%$ of the observations obtain a value less than $z^*$.

```{r}
# P(X <= z^*) = 80%, i.e. how many successes such that 80% of the observations will have that or fewer successes
qbinom(0.8, size = 10, prob = 0.3)
```

```{r}
# P(Z <= z^*) = 95%, i.e. for what value z^* will 95% of the random draws from a standard normal distribution be less than or equal to z^*.
qnorm(0.95, mean = 0, sd = 1)

# What is the 95th percentile of an exam where the mean is 80 and standard deviation is 10
qnorm(0.95, mean = 80, sd = 10)
```











