---
title: "R Homework 3"
subtitle: "Instructions"
author: "Kyle Butts"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: ['default']
    ## No title slide
    seal: false
    self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r preamble, child=here::here("document-preamble.Rmd")}
```


```{r, echo = F}
library(icons)
library(tidyverse)
library(gt)

output_format <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

<h1 class="hi text-2xl center coral">Instructions for R Homework 3*</h1>

<div class="footnote"><sup>*</sup> For all HW assignments, I need to see all the code used</div>


## Question 1

You sample 100 people’s showering habits and find the average shower time is 12.2 minutes. Given that the population variance in shower time is 15, use the `qnorm()` command to construct an 86% confidence interval for a sampled mean. Make sure you are using `qnorm()` correctly. 

.ex[Note:] `qnorm(x)` returns the $Z_x$ such that $P(Z > Z_x) = x$, where $Z \sim N(0,1)$. For a $100 - \gamma$% confidence interval, we have that $\alpha= \frac{1-\gamma}{2}$. See `help(qnorm)` for more details.



## Question 2 

Load the housing data we’ve been using: 

```{r eval = F}
df <- read.csv("https://mattbutner.github.io/data/housing_df.csv")
```

- Use a combination of the `mean()`, `sd()`, `length()`, `qnorm()`, and `sqrt()` functions to construct a 90% confidence interval for the CRIM variables. 

- Interpret the confidence interval from CRIM.
	


---
## Question 3
In the assignment `HW3.Rmd` there is a block of code that simulates a bunch of random samples of the same size, constructs the mean and a confidence interval for each sample, and reports the percent of the confidence intervals that capture the true population mean. 

To begin, make sure you have 
`sample_size <- 100`
`num_samples <- 50`
`ci_level <- 0.95`
set up in the beginning of this chunk. You will need to install the user written package ggplot2. To do this, type `install.packages(“ggplot2”)` into the console before you run the R script. No need to provide the R code for these questions (you can set `include = FALSE`). 

Run the chunk with the initial settings. This will be the comparison.

<ol type = "a">

<li>
Increase the sample size from 100 to 200, by changing `sample_size` and rerunning the code.
<ul>
  <li>What happens to the width of the confidence intervals?</li>
  <li>Does the true population mean fall inside more of the confidence intervals?</li>
</ul>
</li>



<li>
Return the sample size back to 100. Now change the number of samples from 50 to 100.
<ul>
  <li>How does this change the percentage of the confidence intervals that capture the population mean?</li>
  <li>As you increase the number of samples, towards infinity, what percentage of the confidence intervals will capture the true population mean?</li>
</ul>
</li>

<li>
Return the number of samples back to 50. Now change the confidence level to 0.8.
<ul>
  <li>How does the width of the confidence intervals change?</li>
  <li>Does the percentage of confidence intervals that capture the population mean increase or decrease?</li>
</ul>
</li>
</ol>


---
## Question 4 

You started taking the bus to work. The local transit authority says that a bus should arrive at your bus stop every five minutes. After a while, you notice you spend a lot more than five minutes waiting for the bus, so you start to keep a record.

You spend the next two months recording how long it takes for the bus to arrive to the bus stop. This give a total of sixty observations that denote the number of minutes it took for the bus to arrive (rounded to the nearest minute). These observations are hosted at https://mattbutner.github.io/data/bus_stop_time.csv

- Load these data into R as a data frame using the `read.csv()` command.

- Create a histogram of the time_until_bus variable using `hist()`. Would you say that five minutes is a reasonable guess for the average arrival time based on this picture alone?

- Create 95% confidence interval for the bus arrival times using the $z$-distribution. Does 5 minutes fall within the 95% confidence interval?

- How would you communicate your finding to the local transit authority? 




