---
title: "R Homework 1"
subtitle: "Instructions"
author: "Kyle Butts"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: ['default']
    ## No title slide
    seal: false
    self_contained: true
    nature:
      ratio: '8.5:11'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r preamble, child=here::here("document-preamble.Rmd")}
```


```{r, echo = F}
library(icons)
library(tidyverse)
library(gt)

output_format <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

<h1 class="hi text-2xl center coral">Instructions for R Homework 1*</h1>

<div class="footnote"><sup>*</sup> For all HW assignments, I need to see all the code used</div>

This assignment will explore a sample of homes in Boston suburbs. It comes from the paper [Hedonic housing prices and the demand for clean air](sciencedirect.com/science/article/pii/0095069678900062) which tries to estimate how much people are willing to pay to live in homes with cleaner air.


To use this dataset, use the function `read.csv` with the url [https://mattbutner.github.io/data/housing_df.csv](https://mattbutner.github.io/data/housing_df.csv) as done in the introduction_to_R.pdf document. Remember that you need to include the code that loads the dataset into your R Markdown file.

This data set has the following variables

```{r df-info, echo = F}
df_info <- tribble(
  ~Variable, ~Info,
  "CRIM", "Per capita crime rate by town",
  "ZN", "Proportion of residential land zoned for lots over 25,000 sq. ft.",
  "INDUS", "Proportion of non-retail business acres per town",
  "CHAS", "Charles River dummy variable (=1 if census tract touches river; 0 otherwise)",
  "NOX", "Nitric oxides concentration (parts per 10 million)",
  "RM", "Average number of rooms per dwelling",
  "AGE", "Proportion of owner-occupied units built priort to 1940",
  "DIS", "weighted distances to five Boston employment centres",
  "RAD", "index of accessibility to radial highways",
  "TAX", "Full-value property-tax rate per $10,000",
  "PTRATIO", "Pupil-teacher ratio by town",
  "B", "Formula involving % Black",
  "LSTAT", "% lower status of the population",
  "MEDV", "Median value of owner-occupied homes in $1000's"
)

gt(df_info) %>% 
  kfbmisc::gt_theme_kyle() %>% 
  tab_options(table.font.size = px(8))
```



---

## Question 1

Look in the `Environment` panel of R studio, how many variables and how many observations are in this data set? Which of the variables are quantitative, which of the variables are categorical?


## Question 2

Each observation is a town in the suburbs or boston. First, we want to get a sense of the distribution of the *median* value of owner-occupied homes of towns in the Boston suburb at this time. 

Use the function `mean()` and `sd()` to find the average median value of owner-occupied homes in $1000’s. Report the R code and number to two digits (see the option chunk option `echo = TRUE/FALSE` from the [rmarkdown website](https://rmarkdown.rstudio.com/lesson-3.html)).

## Question 3

To practice with R coding. Use the functions `sqrt()`, `sum()`, `^2`, `length()`, and `mean()`, but not `var()` or `sd()`, calculate the sample standard deviation median value of owner-occupied homes in $1000’s. Report the R code and number to two digits.


## Question 4

Use the `hist()` function, create a histogram of NOX pollution. With this distribution is the mean or median a better measure of central tendency?


## Question 5

Now with a sense of the distribution of the two variables of interest. Let's make a scatter plot of `MEDV` on the x-axis and `NOX` on the y-axis. You can use the `plot(x = ____, y = ____)` function for this. Include the plot in the output. To improve the plot, you can use additional parameters `xlab = ""`, and `ylab = ""` to improve the axis labels.

Use the chunk option `fig.cap = ""` to describe the relationship between neighborhood NOX levels and home prices. 





